In this practice, we will use the following data set:

1. Champaign County Sheriff Office & Jail Data

https://data.illinois.gov/organization/champcosheriff

2. 2020 NBA Season Data (stint1,2,3,4)

https://www.basketball-reference.com/leagues/NBA_2021_totals.html


## Champaign County Sheriff Office & Jail Data

```{r}
library(tidyverse)
```
```{r}
jail = read_delim('.../jail_data', delim = ',')
```


Q1. Create a new data object that contains only person booked into jail who resides in Champaign County. Arrange the data object by the city name in alphabetical order and select the following columns: jacket number, age at arrest, city, state, sex, race.

NOTE: 
1. Some of the towns are misspelled, you should correct them.
2. five towns of Champaign county are: Champaign, Danville, Rantoul, Savoy, and Urbana


```{r}
library(sqldf)
```
```{r}
# change date time to character, so they can be fit in sql
jail$booking_time = as.character(jail$booking_time)
jail$released_time = as.character(jail$released_time)
```
```{r}
# build data object cc
cc = sqldf("
      SELECT `jacket_number`, `age arrest`, corrected_city_name AS city, state, sex, race
      FROM(SELECT *, 
      
      /*correct the misspelling for the five towns*/
      
        CASE
          WHEN city = 'CHAMAPIGN' THEN 'CHAMPAIGN'
          WHEN city = 'CHAMPAGIN' THEN 'CHAMPAIGN'
          WHEN city = 'CHAMPIAGN' THEN 'CHAMPAIGN'
          WHEN city = 'CHAAMPAIGN' THEN 'CHAMPAIGN'
          WHEN city = 'CHAMAPIAGN' THEN 'CHAMPAIGN'
          WHEN city = 'CHAMPAIGN,' THEN 'CHAMPAIGN'
          WHEN city = 'CHAMAPAIGN' THEN 'CHAMPAIGN'
          WHEN city = 'CHAMPAIGN' THEN 'CHAMPAIGN'
          WHEN city = 'CHAMAIGN' THEN 'CHAMPAIGN'
          WHEN city = 'CHAMPAING' THEN 'CHAMPAIGN'
        
          WHEN city = 'DANVILLIE' THEN 'DANVILLE'
          WHEN city = 'DAVILLE' THEN 'DANVILLE'
        
          WHEN city = 'SAVOYU' THEN 'SAVOY'
          
          WHEN city = 'RANTOU' THEN 'RANTOUL'
          WHEN city = 'RANOUL' THEN 'RANTOUL'
        
          WHEN city = 'URBNANA' THEN 'URBANA'
          WHEN city = 'URANA' THEN 'URBANA'
        
          ELSE city
          END AS corrected_city_name
        
        FROM jail)
      WHERE city = 'CHAMPAIGN' OR city = 'DANVILLE' 
      OR city = 'RANTOUL' OR city = 'SAVOY' OR city = 'URBANA'
      ORDER BY city
      ")
```
```{r}
# print out 5 rows
sqldf("
      SELECT *
      FROM cc
      LIMIT 10
      ")
```


Q2. How many people booked into jail are recidivists?

NOTE:
Recidivist means a person who goes to jail repeatedly (more than one datetime)

```{r}
sqldf("
      SELECT COUNT(jail_frequency) AS `# of recidivists`
      FROM (SELECT `jacket_number`, COUNT( DISTINCT `booking_date`) AS jail_frequency
      FROM jail
      GROUP BY jacket_number)
      WHERE jail_frequency>1
      ")
```



## 2020 NBA Season Data

```{r}
library('jsonlite')
```
```{r}
stin1 = fromJSON('../NBA_stint1.json')
stin2 = fromJSON('../NBA_stint2.json')
stin3 = fromJSON('../NBA_stint3.json')
stin4 = fromJSON('../NBA_stint4.json')
```

Q1. Reformat the original Player column to PlayerID and PlayerName like the format of other stints data.

EX. 
FROM: {"Player": "Damian Jones\\jonesda03"}
TO:  {"PlayerName": "Damian Jones","PlayerID":"jonesda03"}

```{r}
# separate the first col of stin4 to 2 diff cols
stin4$`PlayerName` = str_split(stin4$Player,"\\\\", simplify = TRUE)[,1]
stin4$`PlayerID` = str_split(stin4$Player,"\\\\", simplify = TRUE)[,2]

# change col order and delete redundant
stin4 = stin4 %>% dplyr::select(!Player) %>% relocate(c(`PlayerID`,`PlayerName`))
```
```{r}
# print 1 st data set
stin1 %>% head(10)
```
```{r}
# print 4 th data set
stin4 %>% head(10)
```


Q2. Summarize the data such that all basketball statistics are totaled across all stints per player, then use the final stint's Position, Age, and Team for each player

```{r}
# combine the four data sets
all_d = sqldf('SELECT * FROM stin1 UNION ALL 
      SELECT * FROM stin2 UNION ALL 
      SELECT * FROM stin3 UNION ALL 
      SELECT * FROM stin4' )
```

```{r}
all_d_s = sqldf("
      SELECT PlayerID, PlayerName, Pos, Age, Tm, stint,
      SUM(G) AS G,
      SUM(GS) AS GS,
      SUM(MP) AS MP,
      SUM(FG) AS FG,
      SUM(FGA) AS FGA,
      SUM(`3P`) AS `3P`,
      SUM(`3PA`) AS `3PA`,
      SUM(`2P`) AS `2P`,
      SUM(`2PA`) AS `2PA`,
      SUM(FT) AS FT,
      SUM(FTA) AS FTA,
      SUM(ORB) AS ORB,
      SUM(DRB) AS DRB,
      SUM(TRB) AS TRB,
      SUM(AST) AS AST,
      SUM(STL) AS STL,
      SUM(BLK) AS BLK,
      SUM(TOV) AS TOV,
      SUM(PF) AS PF,
      SUM(PTS) AS PTS
      FROM all_d
      GROUP BY PlayerID, PlayerNAME
      HAVING stint = MAX(stint)
      ")
```


Q3. Mutate new columns:

1. REBOUNDS: a combination of offensive rebounds and defensive rebounds
2. EFFICIENCY: NBA's efficiency rating using the formula: (points + rebounds + assists + steals + blocks - ((field goals attempted - field goals made) + (free throws attempted - free throws made) + turnovers)) / games played
3. ATO: The assists to turnovers ratio (assists divided by turnovers)
4. PTM: The points to minutes played ratio (points divided by minutes played)
5. 3PP: Three-point shots made out of three-point attempts multiplied by 100
6. de-select the tm and stint columns

```{r}
all_d_s_m = sqldf("
      SELECT `PlayerName`, `PlayerID`, Pos, Age, REBOUNDS,EFFICIENCY
      
      FROM (SELECT *, 
      ORB+DRB AS REBOUNDS, 
      AST/TOV AS ATO, 
      PTS / MP AS PTM,
      (`3P` / `3PA`)*100 AS `3PP`,
      (CAST(PTS+TRB+AST+STL+BLK-((FGA-FG)+(FTA-FT)+TOV) AS float)/CAST(G AS float)) AS EFFICIENCY
      
      FROM all_d_s)
      
      ORDER BY REBOUNDS DESC, `PlayerID`
      ") 
```


Q4. Print the first 10 observations with the columns PlayerName, PlayerID, Pos, Age, Rebounds only.

```{r}
sqldf("
      SELECT `PlayerName`, `PlayerID`, Pos, Age, REBOUNDS
      FROM all_d_s_m
      LIMIT 10
      ")
```


Q5. Which of the 15 most efficient players made the 2020 US Olympic Basketball Team?

NOTE: The member of 2020 US Olympic Basketball Team can be found in the following link.
https://www.teamusa.org/tokyo-2020-olympic-games/meet-team-usa/athletes

```{r}
sqldf("SELECT PlayerName, PlayerID, EFFICIENCY
       FROM (
        SELECT *
        FROM all_d_s_m
        ORDER BY EFFICIENCY DESC
        LIMIT 15
        )
       WHERE PlayerName = 'Bam Adebayo'
       OR PlayerName = 'Devin Booker'
       OR PlayerName = 'Kevin Durant'
       OR PlayerName = 'Jerami Grant'
       OR PlayerName = 'Draymond Green'
       OR PlayerName = 'Jrue Holiday'
       OR PlayerName = 'Keldon Johnson'
       OR PlayerName = 'Zach LaVine'
       OR PlayerName = 'Damian Lillard'
       OR PlayerName = 'JaVale McGee'
       OR PlayerName = 'Khris Middleton'
       OR PlayerName = 'Jayson Tatum'
       OR PlayerName = 'ARIEL ATKINS' 
       OR PlayerName = 'Sue Bird'
       OR PlayerName = 'Tina Charles'
       OR PlayerName = 'Naphesa Collier'
       OR PlayerName = 'Skylar Diggins-Smith'
       OR PlayerName = 'Sylvia Fowles'
       OR PlayerName = 'Chelsea Gray'
       OR PlayerName = 'Brittney Griner'
       OR PlayerName = 'Jewell Loyd'
       OR PlayerName = 'Breanna Stewart'
       OR PlayerName = 'Diana Taurasi'
       OR PlayerName = 'Aja Wilson'
      ")
```