In this practice, we will use the following data:

1. Champaign County Sheriff Office & Jail Data and its data URL

https://data.illinois.gov/organization/champcosheriff

2. Chicago Food Inspections Data

https://data.cityofchicago.org/Health-Human-Services/Food-Inspections-7-1-2018-Present/qizy-d2wf



## Champaign County Sheriff Office & Jail Data

```{r, results='hide', message=FALSE, warning=FALSE}
library(tidyverse)
```
```{r}
champaign_jail = read_delim("../jail.csv",delim = ",")
```
```{r}
# change column format for Jail Data
champaign_jail$booking_date=as.Date(
  champaign_jail$booking_date,  format = '%m/%d/%Y')
champaign_jail$released_date=as.Date(
  champaign_jail$released_date, format = '%m/%d/%Y')
```


Q1. Use two ways to calculate how long does a person stay in jail and compare the result

HINT: 
1. Calculate based on years, days, hours, minutes and seconds
2. Calculate the difference of booking date time and released date time directly

```{r, results='hide', message=FALSE, warning=FALSE}
library(lubridate)
```
```{r}
# create booking date time and released date time
inmates <- champaign_jail %>%
  mutate(booking_date_time=paste0(booking_date," ",booking_time)) %>%
  mutate(released_date_time=paste0(released_date," ",released_time))
inmates$booking_date_time=as.POSIXct(inmates$booking_date_time,
                                     format = '%Y-%m-%d %H:%M:%S')
inmates$released_date_time=as.POSIXct(inmates$released_date_time,
                                      format = '%Y-%m-%d %H:%M:%S')
```
```{r}
# compute how long people spent in jail
df_time=difftime(inmates$released_date_time,inmates$booking_date_time,units = "secs") %>%
  seconds_to_period() %>%
  as.data.frame()
```


### Method 1: Calculate by years, days, hour, minutes, seconds

```{r}
# separate time duration into years, days, hours, minutes and seconds
inmates = inmates%>%
  mutate(years='NA', days='NA', hour='NA', minutes='NA', seconds='NA')

for (i in 1:nrow(inmates)){
  st=str_split(df_time[i,], " ", simplify = TRUE)
  for (j in 1:length(st)){
    if (grepl("Y", st[j], fixed = TRUE)){
      inmates[i,"years"]=str_remove_all(
        str_split(df_time[i,], " ",simplify = TRUE)[j]
        ,"Y+")
    }
    else if (grepl("d", st[j], fixed = TRUE)){
      inmates[i,"days"]=str_remove_all(
        str_split(df_time[i,], " ", simplify = TRUE)[j]
        ,"d+")
    }
    else if (grepl("H", st[j], fixed = TRUE)){
      inmates[i,"hour"]=str_remove_all(
        str_split(df_time[i,], " ", simplify = TRUE)[j]
        ,"H+")
    }
    else if (grepl("M", st[j], fixed = TRUE)){
      inmates[i,"minutes"]=str_remove_all(
        str_split(df_time[i,], " ",simplify =TRUE)[j]
        ,"M+")
    }
    else if (grepl("S", st[j], fixed = TRUE)){
      inmates[i,"seconds"]=str_remove_all(
        str_split(df_time[i,], " ", simplify =TRUE)[j],
        "S+")
    }
  }
}
```
```{r}
# change NA to 0
for (j in c("years", "days", "hour", "minutes", "seconds")){
  inmates[,j]=replace(inmates[,j],inmates[,j]=='NA',toString(0))
}
```
```{r, warning=FALSE}
# change data type
inmates$years <- as.integer(inmates$years)
inmates$days <- as.integer(inmates$days)
inmates$hour <- as.integer(inmates$hour)
inmates$minutes <- as.integer(inmates$minutes)
inmates$seconds <- as.integer(inmates$seconds)

# transform to seconds then transform to hours
for(i in 1:nrow(inmates)){
  inmates$time_spent_in_jail1[i]=round((
    31556926*inmates$years[i]+86400*inmates$days[i]
  +3600*inmates$hour[i]+60*inmates$minutes[i]
  +1*inmates$seconds[i])/3600, digits = 4)
}
```


### Method 2: Calculate by time difference

```{r}
# calculate by the time difference directly
inmates = inmates %>% 
  mutate(time_spent_in_jail2 = 
           round(difftime(inmates$released_date_time,
                    inmates$booking_date_time,units = "hour"), 
                 digit = 4))

inmates$time_spent_in_jail2=
  str_remove_all(inmates$time_spent_in_jail2, " hours$")
inmates$time_spent_in_jail2[is.na(inmates$time_spent_in_jail2)] <- 0
```
```{r, results='hide', message=FALSE, warning=FALSE}
library(data.table)
```
```{r}
# renames columns that contains the word "booking" to "booked"
col_booking <- str_replace_all(colnames(inmates),"booking","booked")
setnames(inmates, old = colnames(inmates), new = col_booking)
```

### Check whether the result of these two methods are the same?

```{r}
result=''
for (i in 1:nrow(inmates)){
  if (inmates$time_spent_in_jail1[i]
      !=inmates$time_spent_in_jail2[i]){
    result='different'
    break
  } else {
    result='The same!'
  }
}
print(result)
```

Q2. For the one that stay in the jail the longest, what crimes does he commit?

```{r}
inmates %>% 
  filter(time_spent_in_jail1 == max(time_spent_in_jail1)) %>% 
  {unique(.$crime)}
```



## Chicago Food Inspections Data

```{r}
chicago_food=read_delim(".../chicago_food.csv",delim = ",")
# correct the date type 
chicago_food$`Inspection Date`=as.Date(chicago_food$`Inspection Date`, format = '%m/%d/%Y')
```

Q1. Create a new data object that meet the following condition:

1. inspection dates is greater than September 01, 2011

2. The data object should contain a column representing whether the inspection is pass or not

3. de-select "Address", "Census Tracts", "City", "Community Areas", "Historical Wards 2003-2015", "Location", "State", "Wards", and "Zip Codes"  

NOTE: 1 if passes, 0 if fails, and NA otherwise

```{r}
inspections <-
  # step1
  chicago_food %>% filter(`Inspection Date`>"2011-09-11") %>%
  # step2
  mutate(pass_or_not = ifelse( Results == "Pass", 1, Results),
         pass_or_not = ifelse( Results == "Fail", 0, pass_or_not),
         pass_or_not = ifelse( Results != "Pass" &Results!="Fail",
                               NA, pass_or_not)) %>%
  # step3
  select(!c("Address", "Census Tracts", "City", 
            "Community Areas", "Historical Wards 2003-2015",
            "Location", "State", "Wards","Zip Codes")) %>%
  # step4
  drop_na()
```


Q2. What was the most common facility type that fail the inspection? 

```{r}
inspections %>% filter(pass_or_not == 0) %>% 
  count(`Facility Type`) %>% 
  arrange(desc(n)) %>% slice(1)
```




